---
title: "Assessment 2 - Birds"
subtitle: "Metropolis-Hastings"
author: "Erik Carrion"
date: "4-20-2023"
output: html_notebook
---

Ornothologists are interested in the ratio of female bird chicks to male bird chicks
From past studies, they know that the probability of a female is distributed Beta(20,20)
If x is the number of female chicks that hatch, then x|p <- Binomial(n, p)
Further, the total hatch size is Poisson(mu = 12).


Construct an MCMC algorithm which will estimate the posterior distributions from p and n
First, derive the posterior conditional distributions analytically
Then, using your analytical results, derive the estimated distributions.

Distributions: 
      p ~ Beta(20,20) = p^19 * (1-p)^19
      x|p ~ Binomial(N, p) = choice(n,x) * (p^x) * (1-p)^(n-x)
      n ~ Poisson(12) = (12^n) * exp(-12) / fact(n)

Posterior Conditional Distributions:
    p|x,n ~ Beta(20 + x, 20 + n - x)
    n|x,p = (12 * (1-p))^(n-x) * exp(-12 * (1-p))/fact(n-x)
```{r}
# Load dependencies
library(tidyverse)
library(gridExtra)
```

 

```{r}
# Set Seed
set.seed(12345, sample.kind = "Rounding")

# Set Number of Iterations
B <- 10^3

# Initialize the model
a0 <- b0 <- 20
lambda <- 12
p <- 1/6
N <- rpois(1,lambda)
X <- 1

# Parameter Storage
p.storage <- X.storage <- N.storage <- numeric(B)

# Start Iterating
for(i in 1:B){
  # Given N, p, X sample a new p
  p <- rbeta(1, a0 + X, b0 + N - X)
  
  # Sample N
  vals <- 6:25
  prob.N <- (12*(1-p))^(vals) / factorial(vals-1)
  N <- sample(6:25, 1, T, prob.N)
  
  # Given the new values of N and P, sample X
  X <- sum(rbinom(N, 1, p))
  
  # Update the monitors
  p.storage[i] <- p
  X.storage[i] <- mean(X)
  N.storage[i] <- N
}
```
``` {r}
# data frame for monitors

df <- data.frame(iter = 1:B,
                 p = p.storage,
                 N = N.storage, 
                 X = X.storage)

# Visualize the distribution of p, n, and X

p.hist <- df %>% ggplot(aes(p)) + geom_histogram() + ggtitle("Hist(p)")
N.hist <- df %>% ggplot(aes(N)) + geom_histogram() + ggtitle("Hist(N)")
X.hist <- df %>% ggplot(aes(X)) + geom_histogram() + ggtitle("Hist(X)")

p.hist; N.hist; X.hist


```



```{r}
summary(df)

```

# Questions: 
  1. What is the most likely estimated egg batch size, N
  2. What is the posterior probability associated with that egg batch size
  3. What is the posterior mean for p
  4. What is the posterior mean for p under a simple beta-binomial with n = 6
  
```{r}

# Q1
mean.N <- mean(N.storage);

# Q2
mean.N.8 <- mean(N.storage == 8);

# Q3
a1 <- a0+1
b1 <- a0+8-1
post.mean.p <- a1/(a1+b1)
post.mean.p;

# Q4
post.mean.p.bb <- (20+1)/(20+1 + 20+6-1)

mean.N;mean.N.8;
post.mean.p;post.mean.p.bb


```
  
  