---
title: "Assessment - Zero Inflated Poisson Model"
output: html_notebook
---
In a national park, 100 tourists were asked how many fish they have caught during their visit. Their 
responses can be found in the data set fish.csv. First we'll load and inspect the distribution of the data. 

```{r}
# Load dependencies & data
library(tidyverse)
library(gridExtra)
fish <- read.csv("fish.csv") %>% filter(is.na(catch)==F)
N = dim(fish)[1]
# Inspect the data
fish %>% ggplot(aes(catch)) + geom_histogram() + ggtitle("Distribution of Caught Fish")
```
We see the number of people responding with zero is significantly larger than 
expected under a Poisson model, so we should be employing a zero-inflated poisson model. 
Before doing so, we'll run a simple Gamma Poisson Model to determine the probability of 
seeing as many zeros as we observed. 

Fit a simple gamma-poisson model and determine the posterior predictive probability p, 
that a sample of 94 people will have at least as many zeros as were observed.

```{r}

# Gamma Posterior Parameters
a.post <- .001+sum(fish$catch)
b.post <- .001+N

post.mean = a.post/b.post
round(post.mean)

# Number of Iterations
B = 10^3

# Set the seed
set.seed(12345, sample.kind="Rounding")

# Sample data from the posterior conditional distribution 
fish.zeros <- replicate(B, {
  # Sample 94 observations using the posterior mean then count & store the # of zeros in the sample
  x <- rpois(94, post.mean)
  sum(x==0)
})
```

What proportion of the observed samples had as many zeros as we observed?
```{r}
# What proportion of our sample saw at least the number of zeros observed in the survey
sum(fish.zeros > sum(fish$catch==0))
```

Given the results of our simulation, the probability of seeing AS many zeros as we did in a sample 
of 94 tourists is p < .001. 

To account for the excess zeros, we now develop a Zero-Inflated-Poisson (ZIP) model which takes 
into account the latent variable, Z, which is 0 or 1 depending on if the person surveyed fished that 
day. 

Because we are modelling excess zeros, we consider p as the probability that Z=0 - i.e. that a rand-
omly surveyed person at the park did NOT fish that day

Priors for Z, p, and mu. 
    Zi ~ Bernoulli(w) // The Joint Distribution of ALL Zi's ~ Binomial(N, w)
    w  ~ Beta(1,1)
    MU ~ Gamma(.001,.001)
    X ~ Poisson(Mu)

```{r}
# Number of iterations
B = 10^4

# monitors for p and mu
pz1.storage <- w.storage <- mu.storage <- numeric(B)
Z.storage <- array(dim = c(94,B))

# Initialize Z, X
X <- fish$catch
Z <- (X>0)*1
```


``` {r}
# Priors for w, mu
# w ~ beta(1,1)
# mu ~ gamma(.001,.001)

# set.seed
set.seed(12345, sample.kind  = "Rounding")
for(i in 1:B){
  # given Z, sample w
  a.beta <- 1 + sum(Z)
  b.beta <- 1 + N - sum(Z)
  
  w <- rbeta(1, a.beta, b.beta)
  
  # Given Z, X sample mu
  a.gamma <- .001 + sum(X)
  b.gamma <- .001 + sum(Z)

  mu <- rgamma(1, a.gamma, b.gamma)
  
  # Given Z,X,& mu sample X
  X <- rpois(N, mu)
  
  # Given Z,X, mu, w sample Z
  pz1 <-  1*mean(X>0) + (exp(-mu)*w)/((exp(-mu)*w) + 1 - w)*mean(X==0)
  Z <- rbinom(N, 1, pz1)
  
  X[Z==0] <- 0
  

  # update monitors
  pz1.storage[i] <- pz1
  w.storage[i] <- w
  mu.storage[i] <- mu
  Z.storage[,i] <- Z
}

```

What do the results of our simulation tell us about the distribution of w, mu, Z, and X?

```{r}
burn <- floor(.20*B)
keep <- B-burn
filter <- rep(c(0,1), c(burn,keep))

print("Summary of Mu - No Burn")
summary(mu.storage)
print("Summary of Mu - After Burning 1st 20% of Iterations")
summary(mu.storage[filter])
print("Summary of W - No Burn")
summary(w.storage)
print("Summary of W - After Burning 1st 20% of Iterations")
summary(w.storage[filter])
print("Summary of P(Z==1|X==0) - No Burn")
summary(pz1.storage)
print("Summary of P(Z==1|X==0) - After Burning 1st 20% of Iterations")
summary(pz1.storage[filter])
```
What do the trace and density plots for mu and w look like?

```{r}
# Set up the data frame and the plots
params.df <- data.frame(iter = 1:B, 
                        MU  = mu.storage,
                        W = w.storage)

trace.mu <- params.df %>% ggplot(aes(iter,MU)) + geom_line() + ggtitle("Trace of MU")
dense.mu <- params.df %>% ggplot(aes(MU)) + geom_density() + ggtitle("Desnity of MU")
trace.w <- params.df %>% ggplot(aes(iter,W)) + geom_line() + ggtitle("Trace of W")
dense.w <- params.df %>% ggplot(aes(W)) + geom_density() + ggtitle("Desnity of W")

my.plots <- list(trace.mu, dense.mu, trace.w, dense.w)
my.layout <- matrix(1:4, nrow=2, ncol=2, byrow = T)

```

```{r}
# Display the plots

grid.arrange(grobs=my.plots,layout_matrix=my.layout, top = "Mu & W Before Burning")
```

Our summary statistics suggest convergence. However, what do the plots look like after we burn them?

```{r}
# Set up the dataframe and plots
burned.df <- data.frame(iter = 801:1000, 
                        MU  = mu.storage[filter],
                        W = w.storage[filter])

trace.mu.b <- burned.df %>% ggplot(aes(iter,MU)) + geom_line() + ggtitle("Trace of MU")
dense.mu.b <- burned.df %>% ggplot(aes(MU)) + geom_density() + ggtitle("Desnity of MU")
trace.w.b <- burned.df %>% ggplot(aes(iter,W)) + geom_line() + ggtitle("Trace of W")
dense.w.b <- burned.df %>% ggplot(aes(W)) + geom_density() + ggtitle("Desnity of W")

my.plots.b <- list(trace.mu.b, dense.mu.b, trace.w.b, dense.w.b)
my.layout.b <- matrix(1:4, nrow=2, ncol=2, byrow = T)
```

```{r}
# Display the plots
grid.arrange(grobs=my.plots.b,layout_matrix=my.layout.b, top = "Mu & W After Burning")
```


```{r}
# Based on Samples
post.mu <- mu.storage[filter] %>% mean()
post.w <- w.storage[filter] %>% mean()

# Posterior predictive distribution of the number of zeros

post.pred.zeros <- replicate(B, {
  Z <- rbinom(N, 1, post.w)
  X <- rpois(N, post.mu)
  X[Z==0] <- 0
  sum(X==0)

})

zeros.df <- data.frame(zeros = post.pred.zeros)

# Distribution of Zeros
zeros.df %>% ggplot(aes(zeros)) + geom_density() + labs(title = "Posterior Distribution for # of Zeros")

```
Now, we ask what is the probability of observing at least as many zeros as observed

```{r}
# Distribution of Zeros using the data
Z.d <- (fish$catch > 0)*1
post.mu.d <- (.001+sum(fish$catch))/(.001+sum(Z.d))
post.w.d <- (1+sum(Z.d))/((1+sum(Z.d)) + 1+94-sum(Z.d))

post.pred.zeros.d <- replicate(B, {
  X <- rpois(N, post.mu.d)
  Z <- rbinom(N, 1, post.w.d)
  X[Z==0]<-0
  sum(X==0)
})

zeros1.df <- data.frame(zeros = post.pred.zeros.d) 
zeros1.df %>% ggplot(aes(zeros))+geom_density()+ggtitle("Predicted Zeros - using Data")

```
```{r}
summary(post.pred.zeros.d); post.mu.d
```
```{r}
# Posterior predictive probability that a sample of 94 will have at least as many zeros as we observed
# 
# 
mean(post.pred.zeros.d>=27)
```

