---
title: "Assessment - Missing Data"
subtitle: "Survey Data"
author: "Erik Carrion"
date: "26 April 2023"
output: html_notebook
---
Suppose we conducted a survey of 100 people. Of those, 10 responded positively, 6 negatively, and 4
didn't respond at all

```{r}
# load dependencies and data
library(tidyverse)
x <- rep(c(1,0,NA), c(40,10,50))
z <- rep(c(1,0), c(50,50))
```

Fit a simple Beta-Binomial model using a uniform prior distribution for the probability of being satisfied. What is the posterior mean probability of being satisfied?


```{r}
# priors for p ~ beta(a,b)
a <- b <- 1

# Set initial values
x.sample <- x;
x.sample[is.na(x)] <- sample(0:1, size=50, replace=T)

# Number of iterations
B <- 10^4 + 500
# monitors for p
p.storage <- numeric(B)

# Gibbs Sampler
for(i in 1:B){
  # Sample p from the posterior conditional distribution
  p <- rbeta(1, a + sum(x.sample), b + sum(1-x.sample))
  
  # Given p, assign the missing value x = 1 with probability p
  x.sample[is.na(x)] <- rbinom(sum(is.na(x)), 1, p)
  
  # Update the monitors
  p.storage[i] <- p
}

# What is the posterior mean probability of being satisfied
mean(p.storage) # yields .6463 which is incorrect
```

What is the posterior central 95% CI for the probability of being satisfied based on the above model? 

```{r}
quantile(p.storage, .975) - quantile(p.storage,.025)
```

Suppose now, that the 50 non responders were all 100% satisfied. What is the posterior mean probability of being satisfied?

```{r}
# priors for p ~ beta(a,b)
a <- b <- 1

# Set initial values
x.sample.2 <- x;
x.sample.2[is.na(x)] <- 1

# Number of iterations
B <- 10^4 + 500
# monitors for p
p.storage.2 <- numeric(B)

# Gibbs Sampler
for(i in 1:B){
  # Sample p from the posterior conditional distribution
  p.2 <- rbeta(1, a + sum(x.sample.2), b + sum(1-x.sample.2))
  
  # Given p, assign the missing value x = 1 with probability 1
  x.sample.2[is.na(x)] <- 1
  
  # Update the monitors
  p.storage.2[i] <- p.2
}

# What is the posterior mean probability of being satisfied
mean(p.storage.2[501:B])
```

# Question 4
Assuming the 50 non responders would have answered in the same way as the 50 responders and employing the same beta-binomial model, what is the width of the posterior central 95% probability of being satisfied?

Given that they didn't respond, we know that the probability of being satisfied is 40/50 = .80 and so we can assign a value of 1 using p = .80

```{r}
# set seed
set.seed(12345,sample.kind="Rounding")
# priors for p ~ beta(a,b)
a <- b <- 1

# Set initial values
x.sample.3 <- x;
x.sample.3[is.na(x)] <- sample(0:1, size=50, replace=T)

# Number of iterations
B <- 10^4 + 500
# monitors for p
p.storage.3 <- numeric(B)

# Gibbs Sampler
for(i in 1:B){
  # Sample p from the posterior conditional distribution
  p.3 <- rbeta(1, a + sum(x.sample.3), b + sum(1-x.sample.3))
  
  # Given p, assign the missing value x = 1 with probability 4p/(3p+1)
  x.sample.3[is.na(x)] <- rbinom(sum(is.na(x)), 1, 4*p/(1 + 3*p))
  
  # Update the monitors
  p.storage.3[i] <- p.3
}

# What is the posterior mean probability of being satisfied
mean(p.storage.3[501:B])
# What is the width of the posterior central 95% probability of being satisfied
quantile(p.storage.3[501:B], .975)-quantile(p.storage.3[501:B], .025)
```



